Deep Reinforcement Learning for Conservation Decisions

### 2021-06-09 appendices/appendix-D-data.Rmd

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(8)
txtcolor <- "#586e75"
```

@@

## Python dependencies loaded via R
sb3         <- import ("stable_baselines3")
gym         <- import ("gym")
gym_fishing <- import("gym_fishing")
```

@@

hake <- fish  %>% filter(commonname=="Argentine hake")
write_csv(hake, "../data/hake.csv")
```

>>>>

@@

## Python dependencies loaded via R
sb3      = import("stable_baselines3")
torch    = import("torch")
gym      = import("gym")
gym_fishing = import("gym_fishing")
gym_conservation = import("gym_conservation")

source("../R/plotting-helpers.R")
```

```{r knitr, include = FALSE}
# This chunk controls plot aesthetics only, and can be omitted with no material change
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())

scale_colour_discrete = function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete = function(...) ggthemes::scale_fill_solarized()
pal = ggthemes::solarized_pal()(8)
txtcolor = "#586e75"

colors <- set_names(c(pal[c(1,3,8,2)], txtcolor, pal[6]),
                    c("Optimal", "TD3_untuned", "TD3", "RL", "historical", "steady-state"))
scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
```

@@

hake <- fish  %>% filter(commonname=="Argentine hake")

hake %>% ggplot(aes(year, scaled_biomass)) + geom_point() + geom_line(aes(year, scaled_catch))
write_csv(hake, "../data/hake.csv")
```

## Applying an RL agent to real data

We use historical data from Argentine Hake to illustrate how an RL agent might be applied in practice.  
Historical biomass and catch data for Argentine Hake can be found in the R.A. Myers Legacy Stock Assessment Database [@ramlegacy].
Here, we load a local copy of the stock assessment for the Argentine Hake from 1986 to 2000.
(Annotated code for data cleaning can be found in Appendix D).

```{r message=FALSE}
hake = read_csv("../data/hake.csv")
td3 = sb3$TD3$load("../python/cache/td3_tuned.zip") 
env = gym$make("fishing-v1")
```

For each year in the data, we compare the historical harvest to the quota that would have been recommended by the TD3-based Deep RL agent we have just trained.
To utilize our RL agent to set a quota for the stock, we only have to map the observed biomass estimates and historical harvests into the re-scaled state space of the environment. 
Recall that the `gym_fishing` environment includes helper function `get_state()` to turn observed biomass to the state-space units used in the gym.

```{r}
# Initial states
x0 = hake[[1, "scaled_biomass"]]
agent_state = agent_action = numeric(15)
agent_state[1] = x0
agent_action[1] = NA
N <- dim(hake)[1]
# `env` is used only to translate between original units and transformed space
# Represent the initial stock size in the 'rescaled' state space:
state = env$get_state(x0)
for(i in 1:N){

  # The RL agent recommends a Quota based on current state
  out = td3$predict(state)
  action = out[[1]]

  # We record state and proposed quota (in the original units)
  agent_state[i] = env$get_fish_population(state)
  agent_action[i] = env$get_quota(action)

  # Use historical data to determine the next state
  state = env$get_state(hake$scaled_biomass[i+1])
}
```

Plot the results:

```{r}
harvest = bind_rows(
  tibble(year = hake$year, harvest = agent_action, model = "TD3"),
  tibble(year = hake$year, harvest = hake$scaled_catch, model = "historical")
)
stock <-  tibble(year =hake$year, biomass = hake$scaled_biomass, model = "historical")

write_csv(harvest, "../manuscript/figs/historical_harvest.csv")
write_csv(stock, "../manuscript/figs/historical_stock.csv")

top_panel <- ggplot(stock, aes(year, biomass, col=model)) + geom_line(lty=2) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
bottom_panel <- ggplot(harvest, aes(year, harvest, col=model)) + 
  geom_line() + geom_point()

top_panel / bottom_panel
```

The Argentine hake biomass exhibited a steep decline between 1986 and 2014 as shown by the stock assessment estimates.
Over that time, historical harvests showed some decline in step with the declining biomass.
By contrast, for each of the historically observed biomass estimates the RL agent would have proposed much steeper cuts to the biomass quota which may have avoided the continued loss of biomass and eventual decline in harvest as well.
This result should not be over-interpreted. 
Without being able to replay history we have no way of truly knowing if the RL agent would have managed this stock better.
While this example shows that it is in principle very straight forward to utilize a trained RL agent to recommend real harvest quotas, we must bear in mind that the agent is only as good as the environments on which it was trained.
In practice, the RL agent would have to first be trained and evaluated on much more realistic environments than only the `fishing-v1` gym illustrated here.

### 2021-06-09 manuscript/manuscript.Rmd

How would an RL agent be applied to empirical data?
In principle, this is straight forward.
Once our agent has been trained on appropriate simulation environments, it is able to propose a quota given an observation of the stock after appropriate transformation of variables.
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ram, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests decreased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.

>>>>

How would an RL agent be applied to empirical data?
In principle, this is straight forward.
After we train an agent on a simulation environment that approximates the fishery of interest, we can query the policy of the agent to find a quota for the observed stock.    
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ram, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests decreased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.
